Approximation as Differentiation: Notes

### 2025-05-20 main/formal-structures.tex
Fwd ref to biproducts:from-product-or-coproduct.

Happily, $\FinVect$ and $\LatGal$ both have biproducts, as a consequence of them having products and being enriched in commutative monoids\todo{fwd ref}. However, neither of them has all small products, without moving to infinite dimensional vector spaces (in the case of $\FinVect$) or complete lattices (in the case of $\LatGal$). Neither of these are ideal. In infinite dimensional vector spaces, dualisation is not involutive and the connection between the forward and reverse derivatives is lost. In complete lattices, we can no longer easily implement the infinite meets and joins required, and we lose the computability of the model. Moreover, in Agda, there are issues with predicativity when considering complete lattices.\todo{be more specific}

@@

\begin{proposition}
\label{prop:biproducts:from-product-or-coproduct}
Suppose $\cat{C}$ is $\CMon$-enriched.
\begin{enumerate}
\item If $(X \times Y, \pi_1, \pi_2)$ is a product then $(X \times Y, \biinj_X, \biinj_Y, \proj_1, \proj_2)$
is a biproduct, where $\biinj_X = \prodM{\id_X}{0_{X,Y}}: X \to X \times Y$ and $\biinj_Y =
\prodM{0_{X,Y}}{\id_Y}: Y \to X \times Y$.
\item If $(X \coprod Y, \inj_1, \inj_2)$ is a coproduct then $(X \coprod Y, \inj_1, \inj_2, \biproj_X,
\biproj_Y)$ is a biproduct, where $\biproj_X: X \coprod Y \to X = \coprodM{\id_X}{0_{Y,X}}$ and $\biproj_Y: X
\coprod Y \to Y = \coprodM{0_{X,Y}}{\id_Y}$.
\item If $X$ is terminal (or initial) in $\cat{C}$ then it is a zero object.
\end{enumerate}
\end{proposition}

>>>>

Happily, $\FinVect$ and $\LatGal$ both have biproducts, as a consequence of their having products and being enriched in commutative monoids~(\propref{biproducts:from-product-or-coproduct} below). However, neither of them has all small products, without moving to infinite dimensional vector spaces (in the case of $\FinVect$) or complete lattices (in the case of $\LatGal$). Neither of these are ideal. In infinite dimensional vector spaces, dualisation is not involutive and the connection between the forward and reverse derivatives is lost. In complete lattices, we can no longer easily implement the infinite meets and joins required, and we lose the computability of the model. Moreover, in Agda, there are issues with predicativity when considering complete lattices.\todo{be more specific}

@@

\begin{proposition}
\label{prop:biproducts:from-product-or-coproduct}
Suppose $\cat{C}$ is $\CMon$-enriched.
\begin{enumerate}
\item If $(X \times Y, \pi_1, \pi_2)$ is a product in $\cat{C}$ then $(X \times Y, \biinj_X, \biinj_Y, \proj_1, \proj_2)$
is a biproduct, where $\biinj_X = \prodM{\id_X}{0_{X,Y}}: X \to X \times Y$ and $\biinj_Y =
\prodM{0_{X,Y}}{\id_Y}: Y \to X \times Y$.
\item If $(X \coprod Y, \inj_1, \inj_2)$ is a coproduct in $\cat{C}$ then $(X \coprod Y, \inj_1, \inj_2,
\biproj_X, \biproj_Y)$ is a biproduct, where $\biproj_X = \coprodM{\id_X}{0_{Y,X}}: X \coprod Y \to X$ and
$\biproj_Y = \coprodM{0_{X,Y}}{\id_Y}: X \coprod Y \to Y$.
\item If $X$ is terminal (or initial) in $\cat{C}$ then it is a zero object.
\end{enumerate}
\end{proposition}
